### 2024-07-29 paper/main.tex
almost done!

\documentclass[12pt,letterpaper]{article}
\usepackage{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{fontspec} %XeLatex
\usepackage{soul,xcolor}
\usepackage{graphicx,wrapfig}
\usepackage{float}
\usepackage{caption}
\usepackage{arydshln} % For \hdashline
\usepackage{lineno}
\usepackage{setspace}

\definecolor{mydarkblue}{RGB}{0,51,102}
\graphicspath{ {./figures/} }

@@

\newcommand{\SuppCaption}[2]{\noindent\textbf{\textcolor{black}{#1}} {#2}\footnotesize}
\newcommand{\SectionHeader}[1]{\noindent\textbf{\Large{\textcolor{black}{#1}}}\normalsize }
\newcommand{\SubSectionHeader}[1]{\noindent\Large{\textcolor{black}{#1}}\normalsize }

@@

    \noindent\textbf{\Large\textcolor{mydarkblue}{Indecisions under time pressure arise from suboptimal switching behaviour
        }}
    \vspace{5mm}
    \\
    Seth R. Sullivan\textsuperscript{1}, \textsuperscript{2} Christopher Peters\textsuperscript{1}, \textsuperscript{2}, Rakshith Lokesh\textsuperscript{1}, Jan A. Calalo\textsuperscript{1},\\ Truc Ngo\textsuperscript{3}, John H. Buggeln\textsuperscript{5},Isaac Kurtzer\textsuperscript{5},Michael Carter\textsuperscript{5}, Joshua G.A. Cashaback\textsuperscript{1},\textsuperscript{2},\textsuperscript{4},\textsuperscript{5}
    \vspace{2mm}
    \\
\end{center}

\textsuperscript{1} Department of Biomedical Engineering, University of Delaware\\
\textsuperscript{2} Department of Mechanical Engineering, University of Delaware\\
\textsuperscript{3} Human Performance Lab, University of Calgary\\
\textsuperscript{4} Biomechanics and Movements Science Program, University of Delaware\\
\textsuperscript{5} Interdisciplinary Neuroscience Graduate Program, University of Delaware\\
\vspace{15mm}
\\
\textcolor{mydarkblue}{\large Abbreviated Title:}
\vspace{1mm}
\\
Indecisive behaviour arises from suboptimal switching
\\
\vspace{15mm}
\\
\textcolor{mydarkblue}{\large Funding:}
\vspace{1mm}
\\
XXXXXXXXXXXXXXX
\\
\vspace{10mm}
\\
\textcolor{mydarkblue}{\large Correspondence:}
\vspace{1mm}
\\
Seth Sullivan\\
Biomedical Engineering\\
University of Delaware\\
STAR Campus, Room 122\\
Newark, DE 19711, U.S.A\\
Email: sethsullivan99@gmail.com
\vspace{2mm}
\\
Or
\vspace{2mm}
\\
Joshua G. A. Cashaback, PhD\\
Biomedical Engineering\\
University of Delaware\\
STAR Campus, Room 201J\\
Newark, DE 19711, U.S.A\\
Email: cashabackjga@gmail.com

\newpage
\doublespace %
\linenumbers %
\noindent \textbf{\large\textcolor{mydarkblue}{ABSTRACT}}
\vspace{2mm}
\\
Indecisive behaviour can be catastrophic, leading to car crashes or stock market losses. Despite fruitful efforts across several decades to understand decision-making, there has been little research on what leads to indecision. Here we examined how indecisions arise under high-pressure deadlines. In our first experiment, participants attempted to select a target by either reacting to a stimulus or guessing, when acting under a high pressure time constraint. We found that participants were suboptimal, displaying a below chance win percentage due to an excessive number of indecisions. Computational modeling suggested that participants were excessively indecisive because they failed to account for a time delay and temporal uncertainty when switching from reacting to guessing, which has not been previously reported in the literature. In a followup experiment, we show for the first time the existence of a time delay and temporal uncertainty when switching from reacting to guessing. Collectively, our results support the idea that humans are suboptimal and fail to account for a time delay and temporal uncertainty when switching from reacting to guessing, leading to indecisive behaviour.
\newpage

\noindent \textbf{\large\textcolor{mydarkblue}{NEW AND NOTEWORTHY}}
\vspace{2mm}
\\
The sensorimotor system has the constant challenge of dealing with the naturally occurring variability in our movements. Here we investigated the potential roles of muscular co-contraction and visuomotor feedback responses to regulate movement variability. When we visually amplified movements, we found that the sensorimotor system primarily uses muscular co-contraction to regulate movement variability. Interestingly, we found that muscular co-contraction was relative to participant-specific visuomotor feedback response, suggesting an interplay between impedance and feedback control.
\newpage
%\noindent\textbf{Author Summary}`
%\vspace{2mm}\\
%
%\newpage
\pagestyle{fancy}
%\cfoot[]{\textcolor{mydarkblue}{\thepage}}
\fancyhead[R]{\emph{\textcolor{mydarkblue}{Indecision arises from suboptimal switching behaviour
        }}}
\fancyfoot[C]{\textcolor{mydarkblue}{\thepage}}
%\fancyfoot[C]{\emph{\textcolor{mydarkblue}{\thepage}}}
%\cfoot{\textcolor{myudblue}{\thepage}}

%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagenumbering{arabic} % page numbering starts here
\noindent\textbf{\large\textcolor{mydarkblue}{INTRODUCTION}}
\vspace{2mm}
\\
Indecisions often arise from failing to decide and act upon sensory information in time, such as a driver failing to brake or hit the gas pedal when a traffic light turns yellow. When acting under high pressure time constraints, the ability to accurately time a decision is critical to success. Past literature has had very little focus on indecisions. The vast majority of decision-making research either does not consider responses made after some time constraint or simply does not permit a non-response, such as in the classic two-alternative forced choice paradigm (Zachsenhouse Robust versus optimal,  Bogacz 2006, Cho et al. 2002 Mechanisms underlying, Jogan and Stocker 2014 A new two-alternative, Ratcliff et al. 2018 modeling 2-alternative forced-choice tasks, Tyler and Chen 2000, Ulrich and Miller 2004 Threshold estimation,   ). Thus, despite its real-world ubiquity and importance, we have very little understanding of how indecisions arise.

There have only been a handful of papers to examine indecisions, which involve either high (Lokesh et al. 2022) or low time pressure (Karsilar et al. 2014, Wu et al. 2016, Phillastides et al. 2011, Dambacher and Hubner 2015, Forstmann et al. 2008). We recently found a high proportion of indecisions during a competitive decision-making task between two humans that observed each other’s movements when selecting a target (Lokesh et al. 2022). In this competitive scenario, the ‘prey’ attempted to end up in the same target as the ‘predator’ by a time constraint, while the predator attempted to end up in the opposite target as the prey. This task had a high time pressure, such that participants were awarded no points if they were indecisive by failing to enter either target within the time constraint. The task poses a conundrum: it may be advantageous to wait for future sensory information and react to an opponent, but it could also be better to switch from reacting to guessing before the time constraint to avoid an indecision. Surprisingly, participants displayed a median indecision rate of approximately 25\% with an upper range close to 40\% indecisions. Similar to others (Wu et al. 2016, Phillastides et al. 2011, Dambacher and Hubner 2015, Forstmann et al. 2008), Karsilar and colleagues (2014) used a low time pressure task and found only 1.7\% of trials were indecisions. Tasks with low time pressure are characterized by providing relatively strong sensory information well in advance of the time constraint deadline. Yet the mechanisms that give rise to indecisive behaviour, which is particularly relevant under common high time pressure scenarios, remains unclear.

Humans and animals attempt to maximize reward to time, select, and indicate a decision with a motor response (drugowitsch, balci, bogacz, hudson landy, trommershauser). To obtain more reward, it has been shown that it is important to consider the inherent time delays and temporal uncertainties of the nervous system (drugowitsch, bogacz, maybe balci too, hudson landy, acerbi wolpert internal representations, Wolpert/Faisal review, Kording/wolpert bayesian decision theory, wolpert/landy motor control is decision-making). Past work has shown that humans will often produce nearly optimal decision times during cognitive (Balci et al. 2011, Miletic and Van Maanen 2019) and motor tasks (Hudson et al. 2008, Faisal and Wolpert 2009 Near Optimal Combination). Other work has shown suboptimal action selection or timing, which has been suggested to occur from an imperfect representation of time delays or temporal uncertainties (Ota et al. 2015, drugowitsch computational precision). With time constraints, misrepresentations of inherent time delays or temporal uncertainties could lead to a missed deadline and consequently an indecision.

Building on our past work (Lokesh, 2022), we developed a high pressure task with a time constraint to examine how humans select decision times. We tested the idea that humans optimally account for time delays and temporal uncertainties to select a decision time that maximizes reward. Alternatively, humans may suboptimally represent time delays and temporal uncertainties, which can lead to an excessive number of indecisions. In Experiment 1, we found humans were suboptimal and observed excessive indecisions that led to a below chance win rate. Computational modelling work suggested that suboptimality arose by failing to account for the time delay and temporal uncertainty associated with switching from reacting to guessing. Experiment 2 showed for the first time, to our knowledge, the existence of an additional time delay and uncertainty when switching from reacting to guessing within a trial. Taken together, our work suggests that humans suboptimally represent the time delay and temporal uncertainty associated with switching from reacting to guessing, leading to indecisive behaviour.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  RESULTS  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\noindent\textbf{\large\textcolor{mydarkblue}{RESULTS}}
\vspace{2mm}
\\
\textbf{\textcolor{mydarkblue}{{Experiment 1}}}
\vspace{2mm}

\noindent\textbf{\textcolor{mydarkblue}{{Experimental Design}}}
\vspace{2mm}

The goal of Experiment 1 was to test how stimulus timing influenced indecisive behaviour. Briefly, participants began each trial by moving their cursor into a start position (Fig. 1A). The stimulus, represented as a cursor on the screen, would quickly move to one of the two target circles. Participants were instructed to reach the same target as the stimulus within a time constraint of 1500 ms. The time remaining in each trial was represented visually with a timing bar that decreased in width according to the elapsed time. Thus, participants were fully aware of how much time they had left relative to the time constraint. A trial was considered a win and the participant received one point if they successfully reached the same target as the stimulus within the time constraint (Fig. 1B). A trial was considered incorrect and the participant received zero points if they reached the opposite target as the stimulus within the time constraint. A trial was considered an indecision and the participant received zero points if they failed to reach a target within the time constraint.

For each trial within a condition, the stimulus movement onset was drawn from the same normal distribution. Using a 3 x 2 within experimental design (Fig. 1C), in separate blocks we manipulated the stimulus movement onset mean (early mean = 1000 ms, middle mean = 1100 ms, late mean = 1200 ms) and standard deviation (low variance = 50 ms, high variance = 150 ms). For the purposes of the main manuscript we focus on the results of the low variance conditions, but report the findings for the high variance conditions in Supplementary A.

\noindent\textbf{\textcolor{mydarkblue}{{Participant Timing Behaviour}}}
\vspace{2mm}

Participant movement onsets for the low variance condition are shown in Fig. 2A, We found a significant main effect of stimulus movement onset mean (F[1.55,29.48] = 4.36, p = 0.030) and variance (F[1.00,19.00], p = 0.017). There was no significant interaction between stimulus movement onset mean and variance (F[1.66,31.47], p = 0.565). When collapsed across low and high variance, participant movement onsets were significantly greater in the middle mean conditions compared to the early mean conditions (p = 0.014, $\hat{\theta}$=72.5 \%), suggesting that participants waited longer to react to the stimulus movement and guessed later in time. Again when collapsed across low and high variance, participant movement onset significantly decreased from the middle mean conditions to the late mean conditions (p = 0.018, $\hat{\theta}$=62.5\%). Here, an earlier participant movement onset in the late mean condition suggests that participants attempted to wait and react to the stimulus, but ended up guessing.

Participant movement onset standard deviation for the low variance conditions are shown in Fig. 2B. There was a main effect of mean (F[1.38, 26.28], p = 0.018) and variance (F[1.00,19.00], p < 0.001) of the stimulus movement onset, and no significant interaction (F[1.98, 37.58], p = 0.097). When collapsed across variance, waiting to react and then guessing in the late mean conditions led to a higher standard deviation of participant movement onset relative to the middle mean (p = 0.039, $\hat{\theta}$ = 70.0\%) and early mean (p < 0.001, $\hat{\theta}$ = 77.5\%) conditions.

\noindent\textbf{\textcolor{mydarkblue}{{Participants are suboptimal and excessively indecisive.}}}

We calculated the indecisions (Fig. 2C), wins (Fig. 2D), and incorrects (Fig. 2E) for each of the experimental conditions. Participants displayed a high proportion of indecisions. The median percentage of indecisions was 15.0\% [range: 0.0 - 93.8\%] across all conditions, with the late mean low variance condition having a median percentage of indecisions of 19.4\% [range: 1.2\%, 93.8\%]. We found a significant interaction between stimulus movement onset and variance for indecisions (F[1.571, 28.781] = 5.58, p = 0.013). In low variance conditions, participants made significantly more indecisions in the middle mean condition than the early mean condition (p<0.001, $\hat{\theta}$ = 85.0\%; Fig. 2C). Additionally, participants made significantly more indecisions in the late mean condition compared to the early mean condition (p<0.004, $\hat{\theta}$ = 80.0\%).

The win percentage across all conditions was 56.25\% (range: 6.2\%, 93.8\%; Fig. 2D]. We found a significant interaction between stimulus movement onset mean and variance for wins (F[1.542, 29.296] = 23.73, p<0.001). We found that the late mean condition had significantly less wins than the early mean condition (p < 0.001, $\hat{\theta}$ = 95.0\%). Interestingly, in the late mean condition we found that the average win percentage was significantly below the 50\% chance level (p<0.001; $\hat{\theta}$=95.0\%), which was true for 19 out of 20 participants. Since guessing on every trial would lead to a win percentage of 50\%, the only way participants would be below chance is if they are excessively indecisive.
The incorrect percentage across all conditions was 26.3\% [range: 0.0\%, 57.5\%; Fig. 2E]. We found significant interactions between stimulus movement onset mean and variance for incorrects (F[1.658,31.508]=3.72, p=0.033). Participants displayed a greater percentage of incorrect decisions in the late mean condition than the early mean condition (p < 0.001, $\hat{\theta}$ = 92.5\%).

\noindent\textbf{\textcolor{mydarkblue}{{Decision making model.}}}

In our task, participants must make a decision of whether to react to the stimulus or guess. For Experiment 1, we tested three different models:  i) No Switch Time Model, ii) Full Switch Time Model and iii) Partial Switch Time Model (Fig. 3, left column). The decision policy of these models considers the expected value ($\mathbb{E[R|\tau]}$) to determine the time ($\tau$) to transition from reacting to guessing. Expected value is defined as

\begin{align}
    \mathbb{E}[R|\tau] = & P(Win|\tau) \cdot R_{Win} \nonumber \\ &+ P(Incorrect|\tau) \cdot R_{Incorrect} \nonumber \\ &+ P(Indecision|\tau) \cdot R_{Indecision}
\end{align}

where ($P(Win|\tau)$) is the probability of a win, ($P(Incorrect|\tau)$) probability of an incorrect, and ($P(Indecision|\tau)$) is the probability of an indecision. $R_{Win} = 1$, $R_{Incorrect} = 0$, and ($R_{Indecision} = 0$ correspond to the reward structure of the task (Fig. 1B). The decision policy of each model maximized expected reward to determine the optimal time to transition from reacting to guessing, $\tau^*$, according to

\begin{equation}
    \tau^* = \underset{\tau}{argmax}[\mathbb{E}(R|\tau)]
\end{equation}

Each model has varying knowledge of the different parameters (Fig. 3, left column). A model can have full knowledge or partial knowledge of a particular parameter. With full knowledge, the decision policy fully utilizes the parameter when selecting the time to transition from reacting to guessing. With partial knowledge, the decision policy utilizes its partial and imperfect representation of the parameter. Here the idea is that a human may not have a perfect representation of some parameter when determining a transition time, even though that particular parameter will still influence behaviour. All model parameter values are shown in Supplementary B.

\noindent\emph{\textcolor{mydarkblue}{{No Switch Time Model}}}

We first considered a model that incorporated various time delays and temporal uncertainties from sources previously identified in the literature: response time, neuromechanical delay, movement time, stimulus movement onset, and timing uncertainty. Note, unlike the other models we will address below, this model did not consider a `switch time’ delay and uncertainty because it was not considered in past literature. Hence, we termed it the No Switch Time Model.

In the late mean low variance condition, the No Switch Time Model underestimated participant movement onset (Fig. 2A), underestimated indecisions (Fig. 2C), and overestimated wins (Fig. 2D). During this condition, participants displayed 19\% indecisions on average and a win percentage significantly below chance. One reason that the No Switch Time Model was unable to capture behaviour is because it did not consider the potential delays and uncertainties that might exist when switching from reacting to guessing.

\noindent\emph{\textcolor{mydarkblue}{{Full Switch Time Model}}}

Next we considered a model that additionally incorporated the potential existence of a switch time delay and uncertainty when transitioning from reacting to guessing. For this Full Switch Time Model, we assumed that the model had full knowledge of the time delay and uncertainty when switching from reacting to guessing.

Yet, despite including switch time, the Full Switch Time Model also performed poorly when predicting participant movement onset (Fig. 2A), indecisions (Fig. 2C), and wins (Fig. 2D) in the late mean condition. An explanation for why this model did not do well to explain indecisions is that humans may not have full knowledge of this potential switch time delay and uncertainty.

>>>>

\documentclass[12pt]{article}
\usepackage{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{fontspec} %XeLatex
\usepackage{soul,xcolor}
\usepackage{graphicx,wrapfig}
\usepackage{float}
\usepackage{caption}
\usepackage{arydshln} % For \hdashline
\usepackage{lineno}
\usepackage{setspace}

\definecolor{mydarkblue}{RGB}{0,51,102}
\definecolor{MYDARKBLUE}{RGB}{0,51,102}

\graphicspath{ {./figures/} }

@@

\newcommand\boldblue[1]{\textcolor{mydarkblue}{\textbf{#1}}}

\newcommand{\SuppCaption}[2]{\noindent\boldblue{{#1}} {#2}\footnotesize}
\newcommand{\SectionHeader}[1]{\noindent\textbf{\Large\textcolor{mydarkblue}{#1}}\normalsize }
\newcommand{\SubSectionHeader}[1]{\noindent\Large{\textcolor{mydarkblue}{#1}}\normalsize }

@@

    \noindent\boldblue{\Large\textcolor{mydarkblue}{Indecisions under time pressure arise from suboptimal switching behaviour
        }}
    \vspace{5mm}
    \\
    Seth R. Sullivan\textsuperscript{1}, Rakshith Lokesh\textsuperscript{4}, Jan A. Calalo\textsuperscript{1,2}, Truc Ngo\textsuperscript{1}, John H. Buggeln\textsuperscript{3},Christopher Peters\textsuperscript{1}, Isaac Kurtzer\textsuperscript{5}, Michael Carter\textsuperscript{6}, Joshua G.A. Cashaback\textsuperscript{1,2,3,7}
    \vspace{2mm}
    \\
\end{center}

\noindent \textsuperscript{1} Department of Biomedical Engineering, University of Delaware\\
\textsuperscript{2} Department of Mechanical Engineering, University of Delaware\\
\textsuperscript{3} Biomechanics and Movements Science Program, University of Delaware\\
\textsuperscript{4} Action Lab, Northeastern University\\
\textsuperscript{5} College of Osteopathic Medicine, New York Institute of Technology\\
\textsuperscript{6} Action, Cognition, and Metascience Lab, McMaster University\\
\textsuperscript{7} Interdisciplinary Neuroscience Graduate Program, University of Delaware\\
\vspace{15mm}
\\
\textcolor{mydarkblue}{\large Abbreviated Title:}
\vspace{1mm}
\\
Indecisive behaviour arises from suboptimal switching
\\
\vspace{15mm}
\\
\textcolor{mydarkblue}{\large Funding:}
\vspace{1mm}
\\
NSF2146888
\\
\vspace{10mm}
\\
\textcolor{mydarkblue}{\large Correspondence:}
\vspace{1mm}
\\
Seth Sullivan\\
Biomedical Engineering\\
University of Delaware\\
STAR Campus, Room 122\\
Newark, DE 19711, U.S.A\\
Email: sethsullivan99@gmail.com
\vspace{2mm}
\\
Or
\vspace{2mm}
\\
Joshua G. A. Cashaback, PhD\\
Biomedical Engineering\\
University of Delaware\\
STAR Campus, Room 201J\\
Newark, DE 19711, U.S.A\\
Email: cashabackjga@gmail.com

\newpage
\doublespace %
\linenumbers %
\SectionHeader{ABSTRACT}
\\
Indecisive behaviour can be catastrophic, leading to car crashes or stock market losses. Despite fruitful efforts across several decades to understand decision-making, there has been little research on what leads to indecision. Here we examined how indecisions arise under high-pressure deadlines. In our first experiment participants attempted to select a target by either reacting to a stimulus or guessing, when acting under a high pressure time constraint. We found that participants were suboptimal, displaying a below chance win percentage due to an excessive number of indecisions. Computational modeling suggested that participants were excessively indecisive because they failed to account for a time delay and temporal uncertainty when switching from reacting to guessing, a phenomenon previously unreported in the literature. In a followup experiment we provide direct evidence for functionally relevant time delay and temporal uncertainty when switching from reacting to guessing. Collectively, our results demonstrate that humans are suboptimal and fail to account for a time delay and temporal uncertainty when switching from reacting to guessing, leading to indecisive behaviour.

\newpage

\pagestyle{fancy}
%\cfoot[]{\textcolor{mydarkblue}{\thepage}}
\fancyhead[R]{\emph{\textcolor{mydarkblue}{Indecision arises from suboptimal switching behaviour
        }}}
\fancyfoot[C]{\textcolor{mydarkblue}{\thepage}}
%\fancyfoot[C]{\emph{\textcolor{mydarkblue}{\thepage}}}
%\cfoot{\textcolor{myudblue}{\thepage}}

\pagenumbering{arabic} % page numbering starts here
\vspace{2mm}
\SectionHeader{INTRODUCTION}
\vspace{2mm}
\\
\noindent Indecisions arise from failing to decide and act upon sensory information in time, such as a driver failing to brake or hit the gas pedal when a traffic light turns yellow. When acting under high pressure time constraints, the ability to accurately time a decision is critical to success. Past literature has had very little focus on indecisions. The vast majority of decision-making research either does not consider responses made after some time constraint or simply does not permit a non-response, such as in the classic two-alternative forced choice paradigm\autocite{zacksenhouse_robust_2010,bogacz_physics_2006,cho_mechanisms_2002,jogan_new_2014,ratcliff_diffusion_2008,ulrich_threshold_2004,tyler_signal_2000}. Thus, despite its real-world ubiquity and importance, we have very little understanding of how indecisions arise.

There have only been a handful of papers to examine indecisions, which involve either high \autocite{lokesh_humans_2022} or low time pressure \autocite{karsilar_speed_2014,wu_capacity_2016,philiastides_causal_2011,dambacher_time_2015,forstmann_striatum_2008}. We recently found a high proportion of indecisions during a competitive decision-making task between two humans that observed each other’s movements when selecting a target \autocite{lokesh_humans_2022}. In this competitive scenario, the ‘prey’ attempted to end up in the same target as the ‘predator’ by a time constraint, while the predator attempted to end up in the opposite target as the prey. This task had a high time pressure, such that participants were awarded no points if they were indecisive by failing to enter either target within the time constraint. The task poses a conundrum: it may be advantageous to wait for future sensory information and react to an opponent, but it could also be better to switch from reacting to guessing before the time constraint to avoid an indecision. Surprisingly, participants displayed a median indecision rate of approximately 25\% with an upper range close to 40\% indecisions. Similar to others \autocite{karsilar_speed_2014,wu_capacity_2016,philiastides_causal_2011,dambacher_time_2015,forstmann_striatum_2008}, Karsilar and colleagues (2014) used a low time pressure task and found only 1.7\% of trials were indecisions. Tasks with low time pressure are characterized by providing relatively strong sensory information well in advance of the time constraint deadline. Yet the mechanisms that give rise to indecisive behaviour, which is particularly relevant under common high time pressure scenarios, remains unclear.

Humans and animals attempt to maximize reward to time, select, and indicate a decision with a motor response \autocite{drugowitsch_optimal_2014,balci_optimal_2011,bogacz_physics_2006,hudson_optimal_2008,trommershauser_humans_2006}. To obtain more reward, it has been shown that it is important to consider the inherent time delays and temporal uncertainties of the nervous system (\autocite{drugowitsch_tuning_2015, balci_optimal_2011,hudson_optimal_2008,acerbi_internal_2012,faisal_noise_2008,kording_bayesian_2006,wolpert_motor_2012,tanis_accuracy_2023,lokesh_visual_2023}. Past work has shown that humans will often produce nearly optimal decision times during cognitive\autocite{balci_optimal_2011,miletic_caution_2019} and motor tasks\autocite{hudson_optimal_2008,faisal_near_2009}. Other work has shown suboptimal action selection or timing, which has been suggested to occur from an imperfect representation of time delays or temporal uncertainties \autocite{ota_motor_2015,drugowitsch_computational_2016}. With time constraints, misrepresentations of inherent time delays or temporal uncertainties could lead to a missed deadline and consequently an indecision. 
Building on our past work\autocite{lokesh_humans_2022}, we developed a high pressure task with a time constraint to examine how humans select decision times. We tested the idea that humans optimally account for time delays and temporal uncertainties to select a decision time that maximizes reward. Alternatively, humans may suboptimally represent time delays and temporal uncertainties, which can lead to an excessive number of indecisions. In \boldblue{Experiment 1}, we found humans were suboptimal and observed excessive indecisions that led to a below chance win rate. Computational modelling work suggested that suboptimality arose by failing to account for the time delay and temporal uncertainty associated with switching from reacting to guessing. \boldblue{Experiment 2} showed for the first time, to our knowledge, the existence of an additional time delay and uncertainty when switching from reacting to guessing within a trial. Taken together, our work suggests that humans suboptimally represent the time delay and temporal uncertainty associated with switching from reacting to guessing, leading to indecisive behaviour.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  RESULTS  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{2mm}
\SectionHeader{RESULTS}

\SubSectionHeader{Experiment 1}

\noindent\boldblue{Experimental Design.}
The goal of \boldblue{Experiment 1} was to test how stimulus timing influenced indecisive behaviour. Briefly, participants began each trial by moving their cursor into a start position (\boldblue{Fig. 1A}). The stimulus, represented as a cursor on the screen, would quickly move to one of the two target circles. Participants were instructed to reach the same target as the stimulus within a time constraint of 1500 ms. The time remaining in each trial was represented visually with a timing bar that decreased in width according to the elapsed time. Thus, participants were fully aware of how much time they had left relative to the time constraint. A trial was considered a win and the participant received one point if they successfully reached the same target as the stimulus within the time constraint (\boldblue{Fig. 1B}). A trial was considered incorrect and the participant received zero points if they reached the opposite target as the stimulus within the time constraint. A trial was considered an indecision and the participant received zero points if they failed to reach a target within the time constraint.

\begin{figure}[H]
  \centering
  \includegraphics[scale =1]{figures/exp1_experimental_design_v2.png}
  \caption{\textbf{Experimental Design.} \textbf{A)} Participants grasped the handle of a robotic manipulandum and made reaching movements in the horizontal plane. An LCD projected images (start position, targets) onto a semi-silvered mirror. Participants began each trial by moving their cursor (purple) into the start position (solid white circle). At the start of the trial (0 ms), they heard a tone and saw two targets (white rings) and a timing bar (white rectangle) appear on the screen. During each trial, a stimulus (yellow cursor) would move quickly in a straight line to one of the two targets. Participants were instructed to reach the same target as the stimulus within a time constraint of 1500ms. This time constraint was visually represented with a timing bar (white rectangle) that decreased in width according to the elapsed time. \textbf{B)} A trial was considered a win and the participant received one point if they successfully reached the same target as the stimulus within the time constraint. A trial was considered incorrect and the participant received zero points if they reached the opposite target as the stimulus within the time constraint. A trial was considered an indecision and the participant received 0 points if they failed to reach a target within the time constraint. \textbf{C)} Stimulus movement onset on each trial was randomly drawn from a specific probability distribution in each condition. Using a within experimental design, we manipulated the mean (early, middle, late) and standard deviation (low, high) of the stimulus movement onset probability distribution. For the main manuscript we focus on results for the low variance conditions, with high variance conditions results shown in \boldblue{Supplementary A.}}
\end{figure}

For each trial within a condition, the stimulus movement onset was drawn from the same normal distribution. Using a 3 x 2 within experimental design (\boldblue{Fig. 1C}), in separate blocks we manipulated the stimulus movement onset mean (early mean = 1000 ms, middle mean = 1100 ms, late mean = 1200 ms) and standard deviation (low variance = 50 ms, high variance = 150 ms). For the purposes of the main manuscript we focus on the results of the low variance conditions, but report the findings for the high variance conditions in \boldblue{Supplementary A}.

\textbf{{{Participant timing behaviour.}}}
Participant movement onsets for the low variance condition are shown in \boldblue{Fig. 2A}, We found a significant main effect of stimulus movement onset mean (F[1.55,29.48] = 4.36, p = 0.030) and variance (F[1.00,19.00], p = 0.017). There was no significant interaction between stimulus movement onset mean and variance (F[1.66,31.47], p = 0.565). When collapsed across low and high variance, participant movement onsets were significantly greater in the middle mean conditions compared to the early mean conditions (p = 0.014, $\hat{\theta}$ = 72.5 \%), suggesting that participants waited longer to react to the stimulus movement and guessed later in time. Again when collapsed across low and high variance, participant movement onset significantly decreased from the middle mean conditions to the late mean conditions (p = 0.018, $\hat{\theta}$ = 62.5\%). Here, an earlier participant movement onset in the late mean condition suggests that participants attempted to wait and react to the stimulus, but ended up guessing.

Participant movement onset standard deviation for the low variance conditions are shown in \boldblue{Fig. 2B}. There was a main effect of mean (F[1.38, 26.28], p = 0.018) and variance (F[1.00,19.00], p < 0.001) of the stimulus movement onset, and no significant interaction (F[1.98, 37.58], p = 0.097). When collapsed across variance, waiting to react and then guessing in the late mean conditions led to a higher standard deviation of participant movement onset relative to the middle mean (p = 0.039, $\hat{\theta}$ = 70.0\%) and early mean (p < 0.001, $\hat{\theta}$ = 77.5\%) conditions.

% Figure 3 Model Outputs and Data
\begin{figure}[H]
    \begin{minipage}[c][\pdfpageheight][t]{0.5\textwidth}

        \includegraphics[scale = 1]{figures/exp1_data_panel_with_models.png}
    \end{minipage}\hfill
    \begin{minipage}[c][\pdfpageheight][t]{0.5\textwidth}
        \caption{
          \textbf{Behaviour Results}: \textbf{A)} Participant movement onset, \textbf{B)} participant movement onset standard deviation, \textbf{C)} indecisions, \textbf{D)} wins, and \textbf{E)} incorrects are shown for each condition. In our high time pressure task, participants made a high proportion of indecisions. Open circles are individual participant data. Filled icons are the predicted behavior of the three models, see legend for detail. Interestingly, we also found that the average win percentage (41.25\%) was significantly below the 50\% chance level (p < 0.001) in the late mean condition, clearly demonstrating suboptimal behaviour. Critically, participants would have earned more points if they had simply guessed earlier on all trials, rather than attempting to react to the stimulus. Model Results: Only the Partial Switch Time Model, which had only a partial representation of the time delay and uncertainty when switching from reacting to guessing, was able to predict the indecisions and a suboptimal win rate (below 50\%) in the late mean condition. See \boldblue{Fig. 3} for the time varying model predictions for the late mean condition. Collectively, our findings show that humans are suboptimal and excessively indecisive.} 
    \end{minipage}
\end{figure}

\textbf{{{Participants are suboptimal and excessively indecisive.}}}
We calculated the indecisions (\boldblue{Fig. 2C}), wins (\boldblue{Fig. 2D}), and incorrect decisions (\boldblue{Fig. 2E}) for each of the experimental conditions. Participants displayed a substantial number of indecisions. The median percentage of indecisions was 15.0\% [range: 0.0 - 93.8\%] across all conditions, with the late mean low variance condition having a median percentage of indecisions of 19.4\% [range: 1.2\%, 93.8\%]. We found a significant interaction between stimulus movement onset and variance for indecisions (F[1.57, 28.78] = 5.58, p = 0.013). In low variance conditions, participants made significantly more indecisions in the middle mean condition than the early mean condition (p<0.001, $\hat{\theta}$ = 85.0\%; \boldblue{Fig. 2C}). Additionally, participants made significantly more indecisions in the late mean condition compared to the early mean condition (p<0.004, $\hat{\theta}$ = 80.0\%).

The win percentage across all conditions was 56.25\% (range: 6.2\%, 93.8\%; \boldblue{Fig. 2D}). We found a significant interaction between stimulus movement onset mean and variance for wins (F[1.54, 29.30] = 23.73, p<0.001). The late mean condition had significantly less wins than the early mean condition (p < 0.001, $\hat{\theta}$ = 95.0\%). Interestingly, in the late mean condition we found that the average win percentage was significantly below the 50\% chance level (p<0.001; $\hat{\theta}$ = 95.0\%), which was true for 19 out of 20 participants. Since guessing on every trial would lead to a win percentage of 50\%, the only way participants would be below chance is if they are excessively indecisive.
The incorrect percentage across all conditions was 26.3\% [range: 0.0\%, 57.5\%; \boldblue{Fig. 2E}]. We found significant interactions between stimulus movement onset mean and variance for incorrect decisions (F[1.66,31.51]=3.72, p=0.033). Participants displayed a greater percentage of incorrect decisions in the late mean condition than the early mean condition (p < 0.001, $\hat{\theta}$ = 92.5\%).

\noindent\textbf{{{Decision-making model.}}}
In our task, participants must make a decision of whether to react to the stimulus or guess. For \boldblue{Experiment 1}, we tested three different models:  i) No Switch Time Model, ii) Full Switch Time Model and iii) Partial Switch Time Model (\boldblue{Fig. 3}, left column). The decision policy of all models considers the expected value ($\mathbb{E}[R|\tau]$) to determine the time ($\tau$) to transition from reacting to guessing. Expected value is defined as

\begin{align}
    \mathbb{E}[R|\tau] = & P(Win|\tau) \cdot R_{Win} \nonumber \\ &+ P(Incorrect|\tau) \cdot R_{Incorrect} \nonumber \\ &+ P(Indecision|\tau) \cdot R_{Indecision},
\end{align}

\noindent where $P(Win|\tau)$ is the probability of a win, $P(Incorrect|\tau)$ is the probability of an incorrect, and $P(Indecision|\tau)$ is the probability of an indecision. $R_{Win} = 1$, $R_{Incorrect} = 0$, and ($R_{Indecision} = 0$ correspond to the reward structure of the task (\boldblue{Fig. 1B}). The decision policy of each model maximized expected reward to determine the optimal time to transition from reacting to guessing ($\tau^*$) according to

\begin{equation}
    \tau^* = \underset{\tau}{argmax}[\mathbb{E}(R|\tau)].
\end{equation}
